---
title: "Heart Disease Data set - An Unsupervised Learning Approach"
author: "Saulo Gil"
date: "06/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r include=FALSE}
library(tidyverse)
library(DT)
library(factoextra)
```

# Pre-processing 
## Loading and adjusting database
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data <- 
  read_csv("processed.cleveland.data", 
    col_names = FALSE)

colnames(data) <- c(
  "age",
  "sex",
  "cp",
  "trestbps",
  "chol",
  "fbs",
  "restecg",
  "thalach",
  "exang",
  "oldpeak",
  "slope",
  "ca",
  "thal",
  "num"
)

data <- 
  data |>
  mutate(ca = as.double(ca),
         thal = as.double(thal)) |> 
  select(-sex)
```

```{r}
datatable(data = data, rownames = FALSE)
```

## Verifying missing values
```{r out.width="100%"}
DataExplorer::plot_missing(data) +
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "none")
```
## Removing missing data
```{r}
data <-
  drop_na(data)
```

## Lets see missing data again
```{r out.width="100%"}
DataExplorer::plot_missing(data) +
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "none")
```

## Data summary
```{r out.width="100%"}
skimr::skim(data)
```
## Scaling data - z-score approach

```{r out.width="100%"}
data_scaled <- 
  scale(data) |> as_tibble()

skimr::skim(data_scaled)
```
# Detecting outliers - Z-score approach (> 2 units)

```{r}

```
# Reducing the dimensionality

```{r out.width = "100%"}
PCA <- princomp(data_scaled)

PCA |> 
  fviz(element = "ind") +
  theme_bw()
```

```{r out.width="100%"}
PCA |>
  plot()
```

# Clustering
## Checking optimal number of clustering using K-Means and elbow method
```{r out.width="100%"}
fviz_nbclust(data_scaled, kmeans, method = "wss")
```

```{r out.width="100%"}
fviz_cluster(kmeans(data_scaled, centers = 2), data = data_scaled) +
  theme_bw()
```


## Checking optimal number of clustering using Hierarchical Clustering and elbow method
```{r out.width="100%"}
fviz_nbclust(data_scaled, hcut, method = "wss")
```

```{r out.width="100%"}
fviz_cluster(hcut(data_scaled, centers = 2), data = data_scaled) +
  theme_bw()

```








